---
title: "10_import"
output:
  html_document:
    df_print: paged
---

# Necessary libraries

```{r}
library(tidyverse)
library(janitor)
library(glue)
library(assertr)
```

# Import data

```{r}
games <- read_csv("Games - Natural Stat TrickTeam Season Totals - Natural Stat Trick (1).csv") %>% 
  clean_names()
```

# Are there any errors in the data?

All but the first three columns of data ("game", "team", and "x3") should be numeric. However, this site reports "NA" values as "-". Columns with a hyphen present appear as character vectors. I need to replace those hyphens with NAs. 

```{r}
# Create function that replaces hyphens with NA
hyphen_na <- function(x) as.double(na_if(x, "-"))

# Apply function to dataset
games <- games %>% 
  mutate_at(vars(toi:attendance), hyphen_na)

# Check to make sure this worked 
summary(games)
```

This dataset records the date and final scores for each game into a single column, which isn't very helpful. I want to split that information into two columns: date and final score. 

```{r}
games <- games %>% 
  separate(col=game, into=c("date", "score"), sep=" - ", remove=TRUE)
head(games)
```

I also want to simply remove the "x3" column, as it contains a classifier from the site that is not useful for us. 

```{r}
games <- games %>% 
  select(-x3)
head(games)
```

One benefit of this dataset is that there should never be a negative value. I need to make sure this is accurate. 

```{r}
games %>% 
  assert(within_bounds(0,Inf),toi:attendance)
```

A regulation NHL game has a duration of sixty minutes. This dataset only includes data from five-on-five situations (in some circumstances, one or both teams will have more or fewer than five players on the ice). Therefore, no "toi" (time on ice) value should be more than 60. I can also assume that no game is ever in non-five-on-five situations more than 50% of the time. I need to make sure this is indeed the case. 

```{r}
games %>% 
  assert(within_bounds(30,60))
```

