---
title: "30_feature_engineering"
output: html_document
---

# Create features to generate and improve model 

## Goal differential

The most important result of a hockey game is, naturally, the final score. As a basic measure of "success", we should go ahead and create a simple integer column for each game's goal differential. Because this dataset only includes data from five-on-five situations and not the entire game, it is possible for a "final goal differential" to be zero. The assumption would be that the other goals contributing to the final score resulted from non-five-on-five situations. 

```{r}
games <- games %>% 
  mutate(differential = gf-ga)
summary(games$differential)
```

## Add some time-related context to values 

This dataset contains exclusively raw values. That is, each non-percentage numeric column represents the total number of that metric produced by the corresponding team in the corresponding game. The "toi" value is context that should be factored into every value, though. It's not very helpful to compare values when one resulted from significantly more time on ice than the other. 

A common way for hockey statisticians to standardize based on TOI is referred to as "per 60" stats. Converting raw values into per 60 is quite simple: 

value per 60 = (raw value/toi) * 60 

Let's go ahead and convert the appropriate columns into per-60 values. 

```{r}
per_60 <- function(x, na.rm=TRUE) ((x/games$toi)*60)

games <- games %>% 
  select(date, score, team, attendance, pdo, toi, contains("percent"), differential, everything()) %>% 
  mutate_at(vars(differential:ldga), list(per_60=per_60))

head(games)
```

## Make save percentage columns more useful

Save percentage and shooting percentage are important values when analyzing hockey. Save percentage (SV%) refers to the number of saves made divided by the number of shots allowed by a team. Typically, these hover around 90% on average. 

SV% = (shots faced - goals allowed) / shots faced 

Shooting percentage (SH%) is exactly the opposite: the number of goals scored divided by shots produced by a team. These average around 10%. 

SH% = goals scored / shots produced

It is often helpful to compare these values. However, you first need to convert save percentage into the following format:

SV% (adjusted) = 100 - SV% (raw)

Let's go ahead and make that adjustment.

```{r}
sv_adj <- function(x, na.rm = TRUE) 100-x 

games <- games %>%
  mutate_at(vars(contains("sv")), list(adj=sv_adj))

head(games)
```

## Make attendance more meaningful 

One interesting variable tracked in this dataset is attendance for each game. Naturally, though, different arenas have different capacities. It is not necessarily informative to compare attendance directly across teams. 

It is somewhat more interesting to compare each game's attendance to that team's season average. 

```{r}
scale1 <- function(x, na.rm=FALSE) (x-mean(x, na.rm=na.rm))/sd(x,na.rm=na.rm)

by_team <- function(df, vars) {
  df %>% 
    group_by(team) %>% 
    mutate_at(vars, list(attendance_std = scale1)) %>% 
    ungroup()
}

games <- games %>% 
  by_team("attendance")

summary(games$attendance_std)
```

# Prepare data for modeling 

The dataset is now ready for modeling. 

```{r}
head(games)
```

There are many options for which variables one would include in a model. A few examples: 

1. The effect of shot quality (high-, mid-, or low-danger) in predicting goals for and against each team. 
2. Is PDO (shooting percentage + save percentage and a common measure of "luck") indeed the most important factor in winning a game? 

3. The effect (if any) of things like attendance and time of year on on-ice performance. 

