---
title: "assign4"
output: html_notebook
---


For the regular expression exercises, try writing first using regular expressions directly, then try using rverbalexpressions. Provide the code for both. 

# Part 1

## R4DS 16.4.5

Create a vector of dates giving the first day of every month in the current year. 

```{r}
library(tidyverse)
library(lubridate)
floor_date(today(), unit="year") + months(0:11)
```

Write a function that given your birthday (as a date), returns how old you are in years.

```{r}
how_old <- function(bday) {
  (bday %--% today()) %/% years(1)
}

how_old("1994-10-19")
```

Write a function that given your birthday, returns the day of the week you were born on. 

```{r}
day_born <- function(date) {
  strftime(date,'%A')
}

day_born("1994-10-19")
```


## R4DS 14.3.2.1

Given the corpus of common words in stringr::words, create regular expressions that find all words that have seven letters or more. (Since this list is long, you might want to use the match argument to str_view() to show only the matching or non-matching words.)

```{r}
library(RVerbalExpressions)

str_view(stringr::words, "\\w{7,}", match = TRUE)

seven <- rx() %>% 
  rx_alpha() %>% 
  rx_multiple(min=7)

str_view(stringr::words, seven, match = TRUE)
```

## R4DS 14.3.3.1

Use str_detect to find all words that:

That only contain consonants. (Hint: thinking about matching “not”-vowels.)

```{r}
no_vowel <- rx() %>% 

str_view(stringr::words, "^[^aeiou]+$", match = TRUE)


```


Empirically verify the rule “i before e except after c”.

```{r}

```


Create a regular expression that will match telephone numbers as commonly written in the US. Split this up into multiple expressions as needed.

```{r}

```

# Part 2


Choose a dataset (this can be the same dataset as the last assignment, or a new one). Define a predictive modeling problem. Create the appropriate files and notebooks to do the following:

1. Describe the data and the problem

2. Read in and check data

2. Clean up the data. 

*Note: You may request that score for these sections replace your score for the coding portion of Assignment 3.*

4. Using best practices, write four functions which add engineered features to the dataset, including a description and rationale. Include a test to determine whether the features are working as intended. 

5. Prepare the data for modeling

Note: this will form the basis for your midterm project. Take advantage of TA's and my office hours. We can provide feedback and guidance. 

